A Comparison of Graph Processing Systems

### 2020-09-24 documentation/paper/sections/results/galoisSpeedup.tex
kleine Korrekturen
Rückmeldung übernommen und todos eingefügt

\paragraph{Time Comparison for Hugepages}
\begin{table}
	\caption{Times for 96 Threads With and Without Hugepages}
	\label{tbl:timeComparison_HPnoHP}
	\renewcommand{\arraystretch}{1.2}
	\centering
	\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
		\toprule
		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
		\cmidrule{3-4}\cmidrule{5-6}
		&\bf Graph&w/o&w/&w/o&w/\\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}
		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5 \\
		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1 \\
		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1 \\
		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4 \\
		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5 \\
		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf BFS}}
		& flickr & 0.01 & \bf 0.00 & 0.6 & \bf 0.3 \\
		& orkut & 0.05 & \bf 0.01 & 0.9 & \bf 0.6 \\
		& wikipedia & 0.52 & \bf 0.12 & 2.4 & \bf 1.4 \\
		& twitter & 7.52 & \bf 3.08 & 14.2 & \bf 6.9 \\
		& rMat27 & 5.62 & \bf 4.79 & 14.6 & \bf 9.2 \\
		& friendster & 3.78 & \bf 1.67 & 12.8 & \bf 7.2 \\
		& rMat28 & 13.14 & \bf 10.25 & 33.1 & \bf 19.1 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Push}}
		& flickr & - & - & 0.6 & \bf 0.3 \\
		& orkut & - & - & 2.1 & \bf 1.2 \\
		& wikipedia & - & - & 13.1 & \bf 11.5 \\
		& twitter & - & - & \bf 74.9 & 201.3 \\
		& rMat27 & - & - & 239.7 & \bf 178.1 \\
		& friendster & - & - & 54.2 & \bf 44.4 \\
		& rMat28 & - & - & 689.3 & \bf 340.1 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Pull}}
		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6 \\
		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4 \\
		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3 \\
		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1 \\
		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1 \\
		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4 \\
		\bottomrule
		\multicolumn{6}{l}{(-) Result is less or equal to 1ms}
	\end{tabular}
\end{table}
When comparing the performance with and without hugepages, the results really speak for themselves they can be seen in \autoref{tbl:timeComparison_HPnoHP}. It is obvious that enabling hugepages greatly improves performance in almost every case. This means that on every algorithm and almost all graphs, calculation time and execution time can be reduced significantly with just the use of hugepages. The only exception is twitter on both PageRank implementations, where the execution time is smaller without hugepages (74.9s vs 201.3s for Push and 8.7s vs 9.3s for Pull with and without hugepages respectively). In all other cases (that's 26 of 28 test cases), the hugepage-version is significantly faster.
The execution time is approximately halved on SSSP and BFS. And on PageRank, the execution time is reduced to 35\% to 87\% of the time without hugepages.

@@

\paragraph{Single-Source Shortest-Paths}
\begin{figure*}
	\hfil
	\begin{subfigure}{0.4\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
		\caption{without hugepages}
		\label{fig:galoisSpeedupSSSP_noHP}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
		\caption{with hugepages}
		\label{fig:galoisSpeedupSSSP_HP}
	\end{subfigure}
	\hfil
	\caption{Calculation time speedups on SSSP}
	\label{fig:galoisSpeedupSSSP}
\end{figure*}
\begin{table}
\renewcommand{\arraystretch}{1.3}
\centering
\caption{Mean Speedups and Variances for SSSP With and Without HugePages}
\label{tbl:ssspMeansVariances}
\begin{tabular}{c
r@{\tabskip 1 \tabcolsep}r
r@{\tabskip 1 \tabcolsep}r}
\toprule
&\multicolumn{2}{c}{\!\!\!$\mu$}&\multicolumn{2}{c}{\!\!$\sigma^2$}\\
\cmidrule(r){2-3}\cmidrule(r){4-5}
{\bf\#Threads}&w/o&w/&w/o&w/\\\midrule
\bf1 & 1.0 & 1.0 & 0.0 & 0.0 \\
\bf2 & 1.6 & 1.6 & 0.2 & \bf 0.1 \\
\bf4 & 2.5 & 2.5 & 1.1 & \bf 0.9 \\
\bf8 & \bf 4.5 & 3.4 & 5.3 & \bf 2.1 \\
\bf16 & \bf 6.7 & 5.8 & 13.0 & \bf 7.3 \\
\bf32 & \bf 9.6 & 7.0 & 38.3 & \bf 10.8 \\
\bf48 & 10.3 & \bf 10.7 & 41.4 & \bf 31.4 \\
\bf96 & 10.2 & \bf 10.8 & 38.1 & \bf 29.9 \\
\bottomrule
\end{tabular}
\end{table}

Starting with SSSP, an algorithm that greatly benefits from many available threads. The results can be seen in \autoref{fig:galoisSpeedupSSSP}.
We first look at the speedups without hugepages, seen in \autoref{fig:galoisSpeedupSSSP_noHP}.
For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads.
Twitter has the best speedup overall. It is 2.6$\times$ with 2 threads compared to one, 4$\times$ with 4, 7.7$\times$ with 8 and 9.7$\times$ using 16 threads.
Behaviour on friendster is similarly good. Here speedup is 1.9$\times$ at 2 threads compared to one, 3.5$\times$ at 4, 6.1$\times$ at 8 threads and 9.7$\times$ at 16 threads.
Anything above 16 threads however no longer helps decrease the computation time significantly on any graph. Speedup above 16 threads is always less than double the speedup of 16 threads. The maximum measured speedups are between 10$\times$ (96 threads, wikipedia) and 19$\times$ (40 threads, rMat28).
In some cases, increasing the thread count can even decrease the speedup again. For example calculation on rMat28 is actually slower with 48 or 96 threads compared to 40 threads. For 40 threads, the speedup is nearly 19$\times$, on 48 threads 17$\times$ and with 96 threads only 15$\times$ compared to one thread.
Small graphs, i.e. flickr and orkut neither benefit from more threads nor is the performance significantly reduced by the synchronization overhead.
Performance on flickr can not be sped up at all, with speedup on flickr being very close to 1 for 1 to 8 threads and between 0.7$\times$ to 0.9$\times$ from 16 to 96 threads.
Orkut reaches maximum speedup of 1.6$\times$ at 16 threads. But on orkut, the speedup is always greater or equal to 1.

@@

\begin{table}
	\caption{Maxium Reached Speedups With And Without Hugepages}
	\label{tbl:maxReachedSpeedupSSSPBFS}
	\renewcommand{\arraystretch}{1.3}
	\centering
	\begin{tabular}{cr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
		\toprule
		&\multicolumn{4}{c}{\bf Max Speedup}\\
		&\multicolumn{2}{c}{\bf SSSP}&\multicolumn{2}{c}{\bf BFS}\\
		\cmidrule{2-3}\cmidrule{4-5}
		\bf Graph & w/o & w/&w/o&w/\\
		\midrule
		flickr & \bf 1.1 &  1.0 & 1.1 & \bf 1.2 \\
		orkut & 1.6 & \bf 4.3 & 1.6 & \bf 5.7 \\
		wikipedia & 10.3 & \bf 15.8 & 4.9 & \bf 10.5 \\
		twitter & \bf 16.8 &  13.0 & 2.5 & \bf 2.7 \\
		rMat27 & 14.0 & \bf 15.8 & \bf 4.2 & 2.9 \\
		friendster & \bf 15.9 &  14.2 & 5.6 & \bf 8.3 \\
		rMat28 & \bf 19.5 &  14.9 & \bf 4.7 & 2.8 \\
		\bottomrule
	\end{tabular}
\end{table}
One might assume the two algorithms SSSP and BFS to behave very similarly, yet they produce very different results.
With SSSP, we observed a decrease in the variance of the speedups, while the average speedup decreased or stayed the same. This provided a more reliable speedup. 
But hugepages did not necessarily increase the maximum possible speedup. For 4 of 7 graphs the larger speedup value is reached without hugepages (cf. \autoref{tbl:maxReachedSpeedupSSSPBFS}).
However exactly this is happening with BFS, we measured a larger speedup with hugepages on all real-world graphs (i.e. 5 of 7).
But as can also be seen in \autoref{fig:galoisSpeedupBFS_HP}, the variance of the speedup increases, making the possible speedup very dependent on the graph.

@@

We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}. Without hugepages, computation time is hardly reduced on any graph other than flickr, where the reached maximum is 1.6$\times$ (cf. \autoref{fig:galoisSpeedupPRPull_noHP}). This maximum is reached at two threads, with speedup steadily declining above that.
The rMat28 is the only other graph of one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 1.3$\times$ at 96 threads.
All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases, and if so only by a small margin.
Computation on Orkut and Twitter reaches a speedup maximum of 1.1$\times$ and 1.05$\times$ at 4 threads, while being less or equal to 1 in all other cases.
Speedup on the wikipedia graph is never grater than one.
Friendster and rMat27 can be sped up to 1.07$\times$ or 1.1$\times$ respectively on 8 threads.

Most of this changes drastically with the activation of hugepages, our data can be seen in \autoref{fig:galoisSpeedupPRPull_HP}. 
Here actually all graphs except flickr reach a speedup greater than 1.5$\times$. Remember, that 1.6$\times$ was the maximum possible speedup without hugepages.
Furthermore, orkut is the only graph that never reaches a speedup of 2$\times$.
Twitter, rMat27, friendster and rMat28 all reach a speedup of around 2.5$\times$ at 96 threads.
Only with hugepages enabled do we observe a considerable increase in performance and thus a benefit of utilizing many threads.
Without hugepages though, having many threads available is in most cases slowing down the performance. Hence, enabling hugepages is very much recommended to be able to use a machine with many cores.

>>>>

\paragraph{Time Comparison for Hugepages}
\begin{table}
	\caption{Times for 96 Threads With and Without Hugepages}
	\label{tbl:timeComparison_HPnoHP}
	\renewcommand{\arraystretch}{1.2}
	\centering
	\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
		\toprule
		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
		\cmidrule{3-4}\cmidrule{5-6}
		&\bf Graph&w/o&w/&w/o&w/\\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}
		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5 \\
		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1 \\
		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1 \\
		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4 \\
		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5 \\
		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf BFS}}
		& flickr & 0.01 & \bf 0.00 & 0.6 & \bf 0.3 \\
		& orkut & 0.05 & \bf 0.01 & 0.9 & \bf 0.6 \\
		& wikipedia & 0.52 & \bf 0.12 & 2.4 & \bf 1.4 \\
		& twitter & 7.52 & \bf 3.08 & 14.2 & \bf 6.9 \\
		& rMat27 & 5.62 & \bf 4.79 & 14.6 & \bf 9.2 \\
		& friendster & 3.78 & \bf 1.67 & 12.8 & \bf 7.2 \\
		& rMat28 & 13.14 & \bf 10.25 & 33.1 & \bf 19.1 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Push}}
		& flickr & - & - & 0.6 & \bf 0.3 \\
		& orkut & - & - & 2.1 & \bf 1.2 \\
		& wikipedia & - & - & 13.1 & \bf 11.5 \\
		& twitter & - & - & \bf 74.9 & 201.3 \\
		& rMat27 & - & - & 239.7 & \bf 178.1 \\
		& friendster & - & - & 54.2 & \bf 44.4 \\
		& rMat28 & - & - & 689.3 & \bf 340.1 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Pull}}
		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6 \\
		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4 \\
		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3 \\
		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1 \\
		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1 \\
		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4 \\
		\bottomrule
		\multicolumn{6}{l}{(-) Result is less or equal to 1ms}
	\end{tabular}
\end{table}
When comparing the performance with and without hugepages, the results show that enabling hugepages greatly improves performance in almost every case, cf. \autoref{tbl:timeComparison_HPnoHP}. This means that on every algorithm and almost all graphs, calculation time and execution time can be reduced significantly with just the use of hugepages. The only exception is twitter on both PageRank implementations, where the execution time is smaller without hugepages (74.9s vs 201.3s for Push and 8.7s vs 9.3s for Pull with and without hugepages respectively). In all other cases (that's 26 of 28 test cases), the hugepage-version is significantly faster.
The execution time is approximately halved on SSSP and BFS. 
On PageRank, the execution time is reduced to a value in the range of 35\% to 87\% of the time without hugepages.

@@

\paragraph{Single-Source Shortest-Paths}
\begin{figure*}
	\hfil
	\begin{subfigure}{0.4\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
		\caption{without hugepages}
		\label{fig:galoisSpeedupSSSP_noHP}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
		\caption{with hugepages}
		\label{fig:galoisSpeedupSSSP_HP}
	\end{subfigure}
	\hfil
	\caption{Calculation time speedups on SSSP}
	\label{fig:galoisSpeedupSSSP}
\end{figure*}
\begin{table}
\renewcommand{\arraystretch}{1.3}
\centering
\caption{Mean Speedups and Variances for SSSP With and Without HugePages}
\label{tbl:ssspMeansVariances}
\begin{tabular}{c
r@{\tabskip 1 \tabcolsep}r
r@{\tabskip 1 \tabcolsep}r}
\toprule
&\multicolumn{2}{c}{\!\!\!$\mu$}&\multicolumn{2}{c}{\!\!$\sigma^2$}\\
\cmidrule(r){2-3}\cmidrule(r){4-5}
{\bf\#Threads}&w/o&w/&w/o&w/\\\midrule
\bf1 & 1.0 & 1.0 & 0.0 & 0.0 \\
\bf2 & \bf 1.6 & 1.6 & 0.2 & \bf 0.1 \\
\bf4 & 2.5 & \bf 2.5 & 1.1 & \bf 0.9 \\
\bf8 & \bf 4.5 & 3.4 & 5.3 & \bf 2.1 \\
\bf16 & \bf 6.7 & 5.8 & 13.0 & \bf 7.3 \\
\bf32 & \bf 9.6 & 7.0 & 38.3 & \bf 10.8 \\
\bf48 & 10.3 & \bf 10.7 & 41.4 & \bf 31.4 \\
\bf96 & 10.2 & \bf 10.8 & 38.1 & \bf 29.9 \\
\bottomrule
\end{tabular}
\end{table}

Starting with SSSP, we observe Galois' SSSP implementation to be greatly benefitting from many available threads. The results can be seen in \autoref{fig:galoisSpeedupSSSP}.
We first look at the speedups without hugepages, seen in \autoref{fig:galoisSpeedupSSSP_noHP}.
For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads.
Twitter has the best speedup overall. It is 2.6$\times$ with 2 threads compared to one, 4$\times$ with 4, 7.7$\times$ with 8 and 9.7$\times$ using 16 threads.
Behaviour on friendster is similarly good. Here speedup is 1.9$\times$ at 2 threads compared to one, 3.5$\times$ at 4, 6.1$\times$ at 8 threads and 9.7$\times$ at 16 threads.
Anything above 16 threads however no longer helps decreasing the computation time significantly on any graph. Speedup above 16 threads is always less than double the speedup of 16 threads. The maximum measured speedups are between 10$\times$ (96 threads, wikipedia) and 19$\times$ (40 threads, rMat28).
In some cases, increasing the thread count can even decrease the speedup again. For example calculation on rMat28 is actually slower with 48 or 96 threads compared to 40 threads. For 40 threads, the speedup is nearly 19$\times$, on 48 threads 17$\times$ and with 96 threads only 15$\times$ compared to one thread.
Small graphs, i.e. flickr and orkut neither benefit from more threads nor is the performance significantly reduced by the synchronization overhead.
Performance on flickr can not be sped up at all, with speedup on flickr being very close to 1 for 1 to 8 threads and between 0.7$\times$ to 0.9$\times$ from 16 to 96 threads.
Orkut reaches maximum speedup of 1.6$\times$ at 16 threads. But on orkut, the speedup is always greater or equal to 1.

@@

\begin{table}
	\caption{Maxium Reached Speedups With And Without Hugepages}
	\label{tbl:maxReachedSpeedupSSSPBFS}
	\renewcommand{\arraystretch}{1.3}
	\centering
	\begin{tabular}{cr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
		\toprule
		&\multicolumn{4}{c}{\bf Max Speedup}\\
		&\multicolumn{2}{c}{\bf SSSP}&\multicolumn{2}{c}{\bf BFS}\\
		\cmidrule{2-3}\cmidrule{4-5}
		\bf Graph & w/o & w/&w/o&w/\\
		\midrule
		flickr & \bf 1.1 &  1.0 & 1.1 & \bf 1.2 \\
		orkut & 1.6 & \bf 4.3 & 1.6 & \bf 5.7 \\
		wikipedia & 10.3 & \bf 15.8 & 4.9 & \bf 10.5 \\
		twitter & \bf 16.8 &  13.0 & 2.5 & \bf 2.7 \\
		rMat27 & 14.0 & \bf 15.8 & \bf 4.2 & 2.9 \\
		friendster & \bf 15.9 &  14.2 & 5.6 & \bf 8.3 \\
		rMat28 & \bf 19.5 &  14.9 & \bf 4.7 & 2.8 \\
		\bottomrule
	\end{tabular}
\end{table}
One might assume the two algorithms SSSP and BFS to behave very similarly, yet they produce very different results. \todo{Begründung, z.b. Frontier Set}
With SSSP, we observed a decrease in the variance of the speedups, while the average speedup decreased or stayed the same. This provided a more reliable speedup. 
But hugepages did not necessarily increase the maximum possible speedup. For 4 of 7 graphs the larger speedup value is reached without hugepages (cf. \autoref{tbl:maxReachedSpeedupSSSPBFS}).
However exactly this is happening with BFS, we measured a larger speedup with hugepages on all real-world graphs (i.e. 5 of 7).
But as can also be seen in \autoref{fig:galoisSpeedupBFS_HP}, the variance of the speedup increases, making the possible speedup very dependent on the graph.

@@

We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}. Without hugepages, computation time is hardly reduced on any graph other than flickr, where the reached maximum is 1.6$\times$ (cf. \autoref{fig:galoisSpeedupPRPull_noHP}). This maximum is reached at two threads, with speedup steadily declining above that.
The rMat28 is the only other graph one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 1.3$\times$ at 96 threads.
All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases, and if so only by a small margin.
% Computation on Orkut and Twitter reaches a speedup maximum of 1.1$\times$ and 1.05$\times$ at 4 threads, while being less or equal to 1 in all other cases.
% Speedup on the wikipedia graph is never grater than one.
% Friendster and rMat27 can be sped up to 1.07$\times$ or 1.1$\times$ respectively on 8 threads.

Most of this changes drastically with the activation of hugepages. Our data can be seen in \autoref{fig:galoisSpeedupPRPull_HP}. 
Here actually all graphs except flickr reach a speedup greater than 1.5$\times$. Remember, that 1.6$\times$ was the maximum possible speedup without hugepages.
Furthermore, orkut is the only graph that never reaches a speedup of 2$\times$.
Twitter, rMat27, friendster and rMat28 all reach a speedup of around 2.5$\times$ at 96 threads.
Only with hugepages enabled do we observe a considerable increase in performance and thus a benefit of utilizing many threads.
Without hugepages though, having many threads available is in most cases slowing down the performance. Hence, enabling hugepages is very much recommended to be able to use a machine with many cores.

### 2020-09-24 documentation/paper/sections/results/pr.tex
some work on discussion/comparison
pagerank introduction
merge und besprechung
Rückmeldung übernommen und todos eingefügt

\subsubsection{PageRank}
\todo{Einleitung}

\paragraph{Single-Node}
Ligra and Polymer support both regular and Delta-PageRank variants.
Ligra's regular PR implementation is faster on 4 of 7 graphs. If the regular version is slower than delta, that is only by a small difference. Explicitly, regular is slower than delta by a range of 6\% to 19\%\ on twitter, rMat27 or friendster. For the other graphs, the delta version is slower by a far greater margin of 13\% to 68\%.
Hence, we only show the results of Ligra's regular PageRank implementation in our evaluation.
For Polymer we found the delta version to be faster on all graphs except rMat28. Delta-PR is on average 15\%\ faster on the first six graphs, while only being 0.3\%\ slower on rMat28.
Thus, the following only shows Polymer's faster Delta-PR implementation.
Giraph required more than the available 250 GB of RAM on any graph larger than wikipedia, hence all of Giraph's results for the larger graphs are missing here.
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
		\caption{Calculation time}
		\label{fig:singleNodePR_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
		\caption{Execution time}
		\label{fig:singleNodePR_exec}
	\end{subfigure}
	\hfil
	\caption{Average times for PR on a single computation node, black bars represent one standard deviation in our testing}
	\label{fig:singleNodePR}
\end{figure*}

The calculation times show some odd behaviour of Galois Push. The required time is less than 1ms, regardless of the graph (cf. \autoref{fig:singleNodePR_calc}). Meanwhile there was no output produced, that would indicate any kind of error. These results would make the calculation times of Galois Push the smallest on all graphs, with a difference of at least one order of magnitude. However, we are very suspicious of these results and thus exclude the calculation time for Galois Push in further comparisons. Because the execution time of Galois Push is always considerably longer than the execution time of Galois Pull (cf. \autoref{fig:singleNodePR_calc}). This leads us to believe that the output that we used for our measurements contains an error.

For the three graphs on which Giraph computed successfully, it is the slowest framework in both calculation and execution times. And that by a difference of three orders of magnitude in the calculation time and one to two orders of magnitude in execution time (cf \autoref{fig:singleNodePR}). On the larger graphs (i.e. those, where there is no data for Giraph), Gemini and Ligra are always slowest in execution time.
Contrary to this, Galois Pull has the smallest execution times on all graphs except flickr (cf. \autoref{fig:singleNodePR_exec}). Ligra is fastest on flickr, while being second fastest on wikipedia, rMat27 and rMat28. Galois Push is second fastest on orkut and friendster.
Interestingly, the execution time for Ligra is at a maximum for twitter. The required time is steadily decreasing with increasing graph size.

\paragraph{Distributed}
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
		\caption{Calculation time}
		\label{fig:distributedPR_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
		\caption{Execution time}
		\label{fig:distributedPR_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTime.png}
		\caption{Overhead}
		\label{fig:distributedPR_overhead}
	\end{subfigure}
	\hfil
	\caption{Average times for PR on the distributed cluster, black bars represent one standard deviation in our testing}
	\label{fig:distributedPR}
\end{figure*}

@@

\todo{Discussion}

>>>>

\subsubsection{PageRank}
In this section we compare the PageRank performance of the various frameworks. As usual we begin with the single-node performance and finish by discussing the distributed variants.

\paragraph{Single-Node}
Ligra and Polymer support both regular and Delta-PageRank variants.
Ligra's regular PR implementation is faster on 4 of 7 graphs. If the regular version is slower than delta, that is only by a small difference. Explicitly, regular is slower than delta by a range of 6\% to 19\%\ on twitter, rMat27 or friendster. For the other graphs, the delta version is slower by a far greater margin of 13\% to 68\%. Because we only run five iterations of PR, the optimizations of the delta algorithm over the regular implementation are not yet visible.
Hence, we only show the results of Ligra's regular PageRank implementation in our evaluation.
Polymer's delta implementation on the other hand can use the optimizations on a much smaller number of iterations to its advantage. For Polymer we found the delta version to be faster on all graphs except rMat28. Delta-PR is on average 15\%\ faster on the first six graphs, while only being 0.3\%\ slower on rMat28. 
Thus, the following only shows Polymer's faster Delta-PR implementation.
Giraph required more than the available 250 GB of RAM on any graph larger than wikipedia, hence all of Giraph's results for the larger graphs are missing here.
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
		\caption{Calculation time}
		\label{fig:singleNodePR_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
		\caption{Execution time}
		\label{fig:singleNodePR_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_overheadTIme.png}
		\caption{Overhead}
		\label{fig:singleNodePR_overhead}
	\end{subfigure}
	\hfil
	\caption{Average times for PR on a single computation node}
	\label{fig:singleNodePR}
\end{figure*}

The calculation times show some odd behaviour of Galois Push. The required time is less than 1ms, regardless of the graph (cf. \autoref{fig:singleNodePR_calc}). Meanwhile there was no output produced, that would indicate any kind of error. These results would make the calculation times of Galois Push the smallest on all graphs, with a difference of at least one order of magnitude. However, we are very suspicious of these results and thus exclude the calculation time for Galois Push in further comparisons. 
We believe that the output we used for our measurements contains an error,
because the \emph{execution} time of Galois Push is always considerably longer than that of Galois Pull (cf. \autoref{fig:singleNodePR_exec}).

For the three graphs on which Giraph computed successfully, it is the slowest framework in both calculation and execution times. Further, it is the slowest by three orders of magnitude in the calculation time and one to two orders of magnitude in the execution time (cf \autoref{fig:singleNodePR}). On the larger graphs (i.e. those, where there is no data for Giraph), Gemini and Ligra are always slowest in execution time.
Contrary to this, Galois Pull has the smallest execution times on all graphs except flickr (cf. \autoref{fig:singleNodePR_exec}). Ligra is fastest on flickr, while being second fastest on wikipedia, rMat27 and rMat28. Galois Push is second fastest on orkut and friendster.
Interestingly, the execution time for Ligra is at a maximum for twitter. The required time is steadily decreasing with increasing graph size.
\todo{Begründung?}

\paragraph{Distributed}
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
		\caption{Calculation time}
		\label{fig:distributedPR_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
		\caption{Execution time}
		\label{fig:distributedPR_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTime.png}
		\caption{Overhead}
		\label{fig:distributedPR_overhead}
	\end{subfigure}
	\hfil
	\caption{Average times for PR on the distributed cluster}
	\label{fig:distributedPR}
\end{figure*}

@@
